---
title: "Exploring the stability of variable selection methods"
author: "Ash Sandhu"
date: "11/29/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glmnet)
library(tableone)
library(caret)
library(ROCR)
library(randomForest)
library(parallel)
library(broom)
library(doParallel)
```


#TODO: 

* Table 1:
  - Cohort demographics
  - Missing values and imputations used
* PreProcessing:
  - Impute missing values
  - Centre and scale everything
  - Model.Matrix
* Methods:
  - StepwiseAIC, univariate-selection based on p-values, LASSO, Elastic-net, RF (Var importance rankings)
*Steps:
  1. Fit model to the entire dataset with a varying sample size 
  (n = 500, 1000, 2000, 3000, max ~ 3500)
  2. Bootstrap entire procedure (1000 times) to get sampling distributions and       frequency of selection

Notes on variables:
temp + temp^2
cont: time_since_hosp, distance_new, #sibllings, maternal_education
Remove: water_source and dataset


##Pre-processing
```{r, echo=FALSE,warning=F,message=F}

#load and reformat data
df.raw <- read_csv('../data/SD_for_Ash.csv')
#change type to categorical for relevant variables
factor.vars <- c("sex","Abnormal_BCS","hivFinal","bednet","final_outcome_death", "boiling", "hivMaternal", "malaria", "sibling_deaths")
del.vars <- c("water_source_drinking")

#clean up
df.raw <- df.raw %>% 
  select(-del.vars) %>% 
  mutate_at(vars(factor.vars), funs(factor))

df.raw$final_outcome_death <- make.names(df.raw$final_outcome_death)
  
table.one <- CreateTableOne(vars = colnames(df.raw), data = df.raw,
                            factorVars = factor.vars)


#centre and scale
df.scaled <- scale(df.raw %>% select(-c(factor.vars, W_AZ, W_LZ,BMI_AZ)))
df.full <- cbind(as.data.frame(df.scaled), df.raw %>% select(factor.vars, W_AZ, W_LZ, BMI_AZ))
df.full <- df.full %>% 
  mutate(temp2 = temperature*temperature)
df.matrix <- model.matrix(final_outcome_death ~ . , df.full)



rm(df.raw)
```


## Model fitting 

##TODO: 
- fix event rate in each sample (~ 5%) -> no need to do this, randomly sampling  should sort this out 
- pull out preformance measures
- add stepwise p-val (not aic)
- change metrics to ROC not accuracy 

###Part 1: Define functions

```{r, warning=F, message=F}

#model fitting procedure
do_model_fitting <- function(df.full,df.matrix, rep.val){

#(a): set up matricies for data collection

model.vars <- colnames(df.matrix)[-1]
n <- c(500,1000,2000,2500,3000,3500,4000)
method <- c("RF", "stepwiseAIC", "univarate.pval", "LASSO")

#variable selection frequency matrix
frequency.matrix <- data.frame(matrix(ncol = ncol(df.matrix) + 1, nrow = 0))
colnames(frequency.matrix) <- c(model.vars,"method", "n")

#helper function to generate samples of size n
get_sample <- function(df, n){
  #outputs a list of dataframes, with samples of varying size
  #method is meant to mimic scenario where data is collected over time so sample size gets larger.
  #n.vec - vector with samples sizes desired
    
  n.vec <- c(n[1],diff(n))
  list.sample <- list(length(n.vec))
  sample.df <- data.frame(matrix(ncol = ncol(df), nrow = 0))
  for(i in 1:length(n.vec)){
    #get sample and index in original df
    sample.df <- rbind(sample.df,sample_n(as.data.frame(df),size = n.vec[i]))
    idx.sample <- plyr::match_df(as.data.frame(df),sample.df) %>% rownames() %>% as.numeric()
    #remove those rows from the old dataframe
    df <- df[-idx.sample, ]
    list.sample[[i]] <- sample.df
  }
  
  return(list.sample)
}

#(b): model fitting methods

#rf method
get_rf_results <- function(sample.list, col.names){
  
  #Get the top variables chosen by RF for different samples sizes 
      
      #helper functions
      fit.rf <- function(df){
        #function that will fit a rf and return top 10 most important varibales
        
        #create model matrix for fitting
        df.matrix <- model.matrix(final_outcome_death ~ . , df)
        #get n and variable names
        n <- nrow(df); model.vars <- colnames(df.matrix)[-1]
        #clean up dataframe to allow for model fitting
        df <- df %>% drop_na()
        #fit rf
        x.mat <- df.matrix[,-1]
        ctrl <- trainControl("cv", 10,classProbs=TRUE, summaryFunction = twoClassSummary )
        df$final_outcome_death <- make.names(df$final_outcome_death)
        rf <- train(x = x.mat, y = df$final_outcome_death,
                    method="rf", metric="ROC", ntree=200,
                    trControl = ctrl, parallel = T)
        #get top variables
        importance <- importance(rf$finalModel)
        top.vars <- row.names((as.data.frame(sort(t(importance)[1,], decreasing = T)[1:10])))
        #score variable with 1 if it is in the model and 0 o.w
        in.model <- as.numeric(model.vars %in% top.vars)
      
        
        return(c(in.model,"RF",n))
        
      }
      
      #helper to speed up
      get_rf_vars <- function(sample.df, names.vec){
        
        #fit row and get vars chosen
        out.row <- as.data.frame(t(fit.rf(sample.df)))
        names(out.row) <- names.vec
        return(out.row)
      
      }
      
      #fit in parallel and get the results
      results <- mclapply(sample.list, get_rf_vars, 
                          names.vec = col.names,
                          mc.cores = 1)
      
      return(results)
}

#stepwise
get_stepwise_results <- function(sample.list, col.names){
  
  
  fit.stepwise <- function(df){
    
    #create model matrix for fitting
    df.matrix <- model.matrix(final_outcome_death ~ . , df)
    #get n and variable names
    n <- nrow(df); model.vars <- colnames(df.matrix)[-1]
    #clean up dataframe to allow for model fitting
    df <- df %>% drop_na()
    #fit stepwiseAIC regression to glm object
    x.mat <- df.matrix[,-1]
    fit.aic <- train(x = x.mat , y = df$final_outcome_death,
                     method = "glmStepAIC",family=binomial(),
                     trControl = trainControl("cv", 10) , trace=F)
    #get vars
    top.vars <- variable.names(fit.aic$finalModel)[-1]
    #score variable with 1 if it is in the model and 0 o.w
    in.model <- as.numeric(model.vars %in% top.vars)

     return(c(in.model,"stepAIC",n))
 
  }


  #helper to speed up
  get_stepaic_vars <- function(sample.df, names.vec){
    
    #fit row and get vars chosen
    out.row <- as.data.frame(t(fit.stepwise(sample.df)))
    names(out.row) <- names.vec
    return(out.row)
  
  }
  
  
  #fit in parallel and get the results
  results <- mclapply(sample.list, get_stepaic_vars,
                        names.vec = col.names,
                        mc.cores = 1)
    
  return(results)

}

#univariate
get_univarate_results <- function(sample.list, col.names){
  
  
  fit.univarate <- function(df){
    
    #create model matrix for fitting
    df.matrix <- model.matrix(final_outcome_death ~ . , df)
    #get n and variable names
    n <- nrow(df); model.vars <- colnames(df.matrix)[-1]
    #clean up dataframe to allow for model fitting
    df <- df %>% drop_na()
    print(table(df$final_outcome_death))
    #fit stepwiseAIC regression to glm object
    x.mat <- df.matrix[,-1]
    ctrl <- trainControl("cv", 10,classProbs=TRUE, summaryFunction = twoClassSummary )
    fit.glm <- train(x = x.mat , y = df$final_outcome_death,
                     method = "glm",family=binomial(),
                     trControl = ctrl,
                     metric = 'AUC', trace=F)
    #get vars
    coef <- tidy(fit.glm$finalModel)
    coef <- coef[-1,]
    top.vars <- coef$term[coef$p.value < 0.05]
    #score variable with 1 if it is in the model and 0 o.w
    in.model <- as.numeric(model.vars %in% top.vars)

     return(c(in.model,"univariate.pval",n))
 
  }


  #helper to speed up
  get_univariate_vars <- function(sample.df, names.vec){
    
    #fit row and get vars chosen
    out.row <- as.data.frame(t(fit.univarate(sample.df)))
    names(out.row) <- names.vec
    return(out.row)
  
  }
  
  
  #fit in parallel and get the results
  results <- mclapply(sample.list, get_univariate_vars,
                        names.vec = col.names,
                        mc.cores = 1)
    
  return(results)

}

#univariate
get_lasso_results <- function(sample.list, col.names){
  
  
  fit.lasso <- function(df){
    
    #create model matrix for fitting
    df.matrix <- model.matrix(final_outcome_death ~ . , df)
    #get n and variable names
    n <- nrow(df); model.vars <- colnames(df.matrix)[-1]
    #clean up dataframe to allow for model fitting
    df <- df %>% drop_na()
    #fit stepwiseAIC regression to glm object
    x.mat <- df.matrix[,-1]
    #fit lasso
    registerDoParallel(4)
    fit.glm <- cv.glmnet(x.mat,df$final_outcome_death, 
                         alpha = 1,family = "binomial",
                         lambda = seq(0.001,0.1,by = 0.001),
                         standardize = F,
                         type.measure = "auc", 
                         nfolds = 10, 
                         intercept = T)
    #get vars
    top.vars <- rownames(coef(fit.glm, s = fit.glm$lambda.min))[coef(fit.glm, s =fit.glm$lambda.min)[,1]!= 0]
    #score variable with 1 if it is in the model and 0 o.w
    in.model <- as.numeric(model.vars %in% top.vars)

     return(c(in.model,"LASSO",n))
 
  }


  #helper to speed up
  get_lasso_vars <- function(sample.df, names.vec){
    
    #fit row and get vars chosen
    out.row <- as.data.frame(t(fit.lasso(sample.df)))
    names(out.row) <- names.vec
    return(out.row)
  
  }
  
  
  #fit in parallel and get the results
  results <- mclapply(sample.list, get_lasso_vars,
                        names.vec = col.names,
                        mc.cores = 1)
    
  return(results)

}

#(c): fit models to samples

#generate samples
sample.list <- get_sample((df.full),n)

#get rf results
rf.results <- get_rf_results(sample.list, colnames(frequency.matrix))
frequency.matrix <- do.call(rbind,rf.results)
print("Finished getting RF results")
#get stepwise results
stepwise.results <- get_stepwise_results(sample.list, colnames(frequency.matrix))
stepwise.results <- do.call(rbind,stepwise.results)
frequency.matrix <- rbind(frequency.matrix,stepwise.results)
print("Finished getting stepwiseAIC results")
#get univariate resutls
univariate.results <- get_univarate_results(sample.list,colnames(frequency.matrix))
univariate.results <- do.call(rbind,univariate.results)
frequency.matrix <- rbind(frequency.matrix, univariate.results)
print("Finished getting univariate filtering results")
#get LASSO resutls
lasso.results <- get_lasso_results(sample.list,colnames(frequency.matrix))
lasso.results <- do.call(rbind,lasso.results)
frequency.matrix <- rbind(frequency.matrix, lasso.results)
print("Finished getting LASSO results")
frequency.matrix$rep_val <- rep.val

return(frequency.matrix)
}

```


###Part 2: Running the porcedure 

```{r}
out <- list(10)
for(i in 1:10){
  out[[i]] <- do_model_fitting(df.full,df.matrix,i)
}

out <- do.call(rbind,out)
```


